\section{Entities and Terminology}

In this white paper we use the following terminology for the entities in a presence-tracing system. All presence tracing systems we examine, whether digital or analog, use the following high-level entities or a subset thereof:

\begin{description}[topsep=2pt, partopsep=2pt]
\item[Visitor] somebody who visits a location. Sometimes called user. There can be many visitors.
\item[Location] a semi-public location (for example, bar, restaurant, religious building, events venue, or meeting place) or an event which takes place in one or more locations (for example, an exhibition). There can be many locations.
\item[Location Owner] the owner or manager of a location. For simplicity, we assume each location has one owner. An owner may manage many locations.
\item[Health Authority] the public health-authority that determines which visitors of which locations to notify. Usually, there is only one health authority.
\item[Backend] a backend server that can (potentially) interact with visitors, locations, and the health authority. Usually there is only one backend.
\end{description}

Presence tracing concerns crowds that gather at a certain location at a certain time. This could be a crowd on a Friday evening at a bar, or on a Tuesday afternoon meeting in a community center. For readability we use the term ``location'' to refer to locations at a specific time as well as specific events (that may happen within one or more physical locations).

Health authorities may use different criteria to determine when to notify visitors to a location for a given time slot  (e.g., visitors were wearing masks, there were barriers or large separation between groups). We use the following terminology to indicate locations whose visitors should be notified and the respective visitors:
\begin{description}[topsep=2pt, partopsep=0pt]
\item[Trace location] a location that was visited by one or more SARS-CoV-2-positive persons, and whose visitors should be notified.
\item[Presence contact] a person that was present at a location around the same time as one or more SARS-CoV-2-positive persons and should thus be notified.
\end{description}
